---
title: "Gun_control_chi_square"
author: "Andrew Paterson"
date: "November 13, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This: I'm andrew and I'm updating this markdown file here.

```{r, echo = FALSE}

library(tidyverse)

gc_df <- read_csv("gc_df.csv")

gc_df

```

We are asking: Is there a significant difference in recommended gun control action for male and female correspondents, where the question is, should guns be easier, harder, or the same difficulty to get?

Hey Andrew love this question - real insightful 

Thanks Sara, you're a real team player, and I value your contributions to our project.

Andrew data wrangling and counts

```{r}
# Create data into contingency form that works for chi square 
get_guns <- gc_df %>% 
  filter(gender != "NA", make_get_guns != "NA") %>% 
  count(gender, make_get_guns) %>% 
  spread(make_get_guns, n) %>% 
  select(-gender)

rownames(get_guns) <- c("Female", "Male")

```

Sara the collaborating QUEEN is about to do some cool data stuff.... like running a Chi-square test for independence

We ask: "Is there a significant association between sex and opinion on whether guns should be more difficult, easier, or the same difficulty to get?"

```{r}

# Let's look at the actual proportions:

guns_prop <- prop.table(as.matrix(get_guns), 1) # Margin for calculating proportions (the 1 tells R to calculate proportions by rows)


# Run chi-square test for independence 
# H0: No significant difference in proportions for male and female respondents 

guns_x2 <- chisq.test(get_guns)
guns_x2

# Test for drivers of significance: in the console use guns_x2stdres. Anything over absolute value of 2 could be considered a driver of significance. 
```
"Recommended government action on gun access differed significantly between male (n = ) and female (n = ) respondents ($\chi^2$(`r guns_x2$parameter`) = ....,  p < 0.001. Notably, a much higher proportion of female respondents (57.4%) recommend making it more difficult to get guns, compare to 40.4% for male respondents." 
